Open-Source Web-Based Visualizer for \\ Dynamic-Response Shared Taxi Simulations

### 2020-12-30 src/paper.tex
submitted version 30-Dec-2020

%% The `ecrc' package must be called to make the CRC functionality available
\usepackage{ecrc}
\usepackage[bookmarks=false]{hyperref}
    \hypersetup{colorlinks,
      linkcolor=blue,
      citecolor=blue,
      urlcolor=blue}
%\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}

@@

We describe a unique, web-based data visualization portal developed for use by researchers and public transit agencies investigating future shared-taxi fleet scenarios. Augmenting or even replacing fixed-route transit lines with automated, connected, shared taxi fleets may be a desirable alternative in less-densely developed areas. The MATSim agent-based transport microsimulation model is used to study scenarios including status quo; dynamically-dispatched fleets with drivers; and fully autonomous fleets. This paper focuses on a data visualization portal which includes many interactive views, such as agent (taxi) movements color-coded by number of passengers and trip request origins and destinations, changes in roadway and passenger volumes compared to a base case, and more. The agent-based simulation covers a 24 hour simulation period; analysts can hone in on specific times of day to examine, e.g. school pickup/drop-offs or commute trips connecting to rail stations. The tool is in operation for several small cities and rural regions in Germany and was successfully used as an outreach tool in public meetings. In addition, developers of the MATSim DRT extension found the visualizations particularly useful for debugging both the algorithms and the scenario definitions. All code is open source and, while this specific study has a very specific use case, the portal has an extensible design that could be modified for other purposes.

@@

\begin{keyword}
shared taxis; data visualization; matsim; public outreach
\end{keyword}

@@

The existing methodology behind using MATSim to simulate DRT is well-described by [*] and is not the focus of this paper. Rather, in the course of the effort to develop, analyze and compare various DRT scenarios, a need arose to create some truly unique data visualizations that could assist in improving and displaying the DRT scenarios themselves. A web-based portal is the result of this research. The portal includes many interactive views, such as agent (taxi) movements color-coded by number of passengers and trip request origins and destinations, changes in roadway and passenger volumes compared to a base case, and more. The agent-based simulation covers a 24 hour simulation period; analysts can hone in on specific times of day to examine, e.g. school pickup/drop-offs or commute trips connecting to rail stations.

The tool is now in operation for several small cities and rural regions in Germany, and was successfully used as an outreach tool in online public meetings. Project partners at transit agencies are interested in understanding the detailed implications of the DRT simulations, and without a visual, interactive component, that would be extremely difficult to provide.

@@

Agent-based microsimulation is quite complex and the reams of data and overwhelming details are not appropriate for public outreach and stakeholder meetings. PDF reports and PowerPoint slides are the norm in these situations, but there is an opportunity here to build something truly interactive, possibly an order of magnitude more useful.

Many existing tools can analyze agent-based transport simulation outputs. \textbf{OTFVis} is a venerable open-source tool designed to integrate tightly with MATSim \cite{Srippgen2015OTFVisInBook}, however it is not under active development for several years and has no online component nor DRT-specific capabilities.

@@

General geographic information system tools, such as \textbf{ArcGis} and its open source corollary \textbf{QGis}, are again very analyst-focused. Unlike QGis, ArcGis can create high quality online dashboards and interactive maps, and study team has used them in other settings. ArcGis does not have any MATSim-specific capabilities; the rich agent-based details would be entirely lost in the aggregations required to get DRT outputs onto an ArcGis map. ArcGis is also commercial software and was not considered due to the required financial outlay.

Research by \citet{CharltonLaudan2020WebBasedVisualization} shows that a web-based tool specific to MATSim is now possible, given the latest advances in 3D technology embedded in modern web browsers. That tool, known as "MatHub", includes a web front-end (website) which runs in the user's browser. Through this site the user organizes model outputs, and defines aggregate visualizations for a few typical use cases. The tool also requires a large, complex back-end server component which handles file storage, user authentication, output post-processing, and realtime visualization support. That work is an obvious starting point for this research, but for reasons that will be described below, much of it was ultimately discarded.

@@

Despite two years of development of the previous client/server "MatHub" solution for other studies, many problems remained. The two largest were: (1) The massive size of MATSim outputs (a single run can be 800Mb or more) always preclude loading them into a user's web browser. This forces post-processing and/or aggregation to happen, which can eliminate many of the benefits of a fully disaggregate analysis. (2) The back-end server design required extensive effort whenever new visualization capabilities were desired. A strict database design and overwrought user/group management capabilities resulted in a rock-solid system that was extremely onerous to modify. It turns out that complex file and user management was unnecessary.\footnote{This invokes the programming mantra "YAGNI" -- You Aren't Gonna Need It: just build the simplest possible solution that works instead of over-engineering things.}

Trial and error eventually converges on a simplified system that eliminates all of the back-end server complexity by eliminating all of the back-end servers:

\begin{itemize}
  \item File storage is delegated to existing online storage, instead of creating a special visualization server. On site, we have an existing "Subversion" server \cite{Collins-SussmanEtc2008SubversionBook} which is quite adept at large-file storage and version management. All access to the server is via a very simple interface that can be easily extended to cloud storage, such as Amazon S3
  \item As the site is intended to be publicly accessible, we discard all user authentication capabilities. If you have a web browser, you can view the results of any runs which are stored \emph{in the public area} of the Subversion server. Naturally some parts of file storage remain off-limits to the greater Internet
  \item Analysts define visualizations by creating small configuration text files in the same folder as model outputs. These files include all of the details of a visualization --- usually just a few lines --- pointing to the input files and configuration settings. This eliminates all the hassle of writing code for user interface input dialog boxes.
  \item Post-processing of outputs must be performed by the analyst as part of their modeling workflow, before uploading to Subversion, since browsers cannot ingest MATSim outputs in their entirety.
\end{itemize}

@@

\begin{itemize}
  \item The overall application framework is based on \textbf{VueJS}, a popular Javascript toolkit for building interactive websites
  \item The agent-based visualizations use custom code ("shaders") written using the \textbf{Deck.gl} visualization library
  \item Older visualizations borrowed from Charlton's previous research use the \textbf{MapBox GL} mapping library
  \item Extensive optimization of memory usage for handling these large datasets with the brilliant \textbf{crossfilter} Javascript library, which manages lightning-fast filtering of million-rows datasets in multiple dimensions
\end{itemize}

The entire codebase is open source and is designed with a "plug-in" architecture so that new visualizations can be developed quickly while relying on common infrastructure for things like file access.

To create a new visualization, the developer copies an existing template, modifies the Javascript code to take advantage of the Deck.gl or MapBox libraries as needed, and registers it with the main application. End users can then create visualizations by adding corresponding configuration files to a scenario's file server folder.

@@

These descriptions fit in this scientific paper but for best results it is easiest to view the site in situ, as it remains online and is better demonstrated than described.\footnote{Site is viewable at avoev-vsp.github.io}

@@

Another aggregate visualization of total mode shift from one scenario to another, in this study comparing DRT scenarios to the base case. These visualizations really highlight the degree to which the new DRT mode is "stealing" trips from existing transit and bicycle modes versus from private vehicles.

@@

Throughout summer 2020, our team iterated on dozens of DRT scenarios: varying pricing structures, presence of drivers, service coverage areas, maximum wait time goals, and so on. In August 2020 we began the first public meetings with transit agency staff, with the fully operational website at \url{https://avoev-vsp.github.io} (available in German only).

@@

Several years of trial and error led us to this specific combination of technologies and techniques. The research team is deeply indebted to the developers of so many components and libraries on which this is based, all of which are freely available online and without which this work would have been entirely impossible. The final product is more than parts merely cobbled together: a usable, useful tool for decision-making now exists, and the DRT-specific extensions serve as a template for other work.

@@

The website remains online at \url{https://avoev-vsp.github.io}. All code is available at \url{https://github.com/avoev-vsp/avoev-vsp.github.io} and is licensed under the GNU GPL Version 3 \cite{FSF2007GnuGPL}. The authors welcome feedback and code contributions.

>>>>

%% The `ecrc' package must be called to make the CRC functionality available
\usepackage{ecrc}
\usepackage[bookmarks=false]{hyperref}
    \hypersetup{colorlinks,
      linkcolor=blue,
      citecolor=blue,
      urlcolor=blue}
%\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[utf8]{inputenc}
% make sure quotation marks look correct
\usepackage [english]{babel}
\usepackage [autostyle, english = american]{csquotes}
\MakeOuterQuote{"}

@@

We describe a unique, web-based data visualization portal developed for use by researchers and public transit agencies investigating future shared-taxi fleet scenarios. Augmenting or even replacing fixed-route transit lines with automated, connected, shared taxi fleets may be a desirable alternative in less-densely developed areas. The MATSim agent-based transport microsimulation model is used to study scenarios including the status quo, dynamically-dispatched fleets with drivers, and fully autonomous fleets. This paper focuses on a data visualization portal which includes many interactive views such as agent (taxi) movements color-coded by number of passengers and trip request origins and destinations, changes in roadway and passenger volumes compared to a base case, and more. The agent-based simulation covers a 24 hour simulation period; analysts can hone in on specific times of day to examine, e.g. school pickup/drop-offs or commute trips connecting to rail stations. The tool is in operation for several small cities and rural regions in Germany and was successfully used as an outreach tool in public meetings. In addition, developers of the MATSim DRT extension found the visualizations particularly useful for debugging both the algorithms and the scenario definitions. The code is entirely open source and, while this specific study has a rather esoteric use case, the visualization platform has an extensible design that could be modified for other purposes.

@@

\begin{keyword}
dynamic response transit; data visualization; agent based simulation; matsim; shared taxis
\end{keyword}

@@

The existing methodology behind using MATSim to simulate DRT is well-described by Bischoff et al \cite{Bischoff2020,BischoffMaciejewskiNagel2017SharedTaxiIITSC} and is not the focus of this paper. Rather, in the course of the effort to develop, analyze and compare various DRT scenarios, a need arose to create some truly unique data visualizations that could assist in improving and displaying the DRT scenarios themselves. A web-based portal is the result of this research. The portal includes many interactive views, such as taxi movements color-coded by number of passengers and trip request origins and destinations, changes in roadway and passenger volumes compared to a base case, and more. The agent-based simulation covers a 24 hour simulation period; analysts can hone in on specific times of day to examine, e.g. school pickup/drop-offs or commute trips connecting to rail stations.

The tool is in operation for several small cities and rural regions in Germany, and was successfully used as an outreach tool in public meetings online. Project partners at transit agencies are interested in understanding the implications of the DRT simulations, and without a visual component, that would be extremely difficult to provide.

@@

Agent-based microsimulation is quite complex, producing reams of data and overwhelming details that are not appropriate for stakeholder meetings. PDF reports and PowerPoint slides are the norm in these situations, but building something truly interactive instead could be much more engaging and useful for decision-making.

Many existing tools can analyze agent-based transport simulation outputs. \textbf{OTFVis} \cite{Srippgen2015OTFVisInBook} is a venerable open-source tool designed to integrate tightly with MATSim, however it is not under active development and has neither online features nor DRT-specific capabilities.

@@

General geographic information system tools, such as \textbf{ArcGis} and its open source corollary \textbf{QGis}, are again very analyst-focused. Unlike QGis, ArcGis can create high quality online dashboards and interactive maps, and the study team has used them in other settings. ArcGis does not have any MATSim-specific capabilities; the rich agent-based details would be entirely lost in the aggregations required to get DRT outputs onto an ArcGis map. ArcGis is also commercial software and was not considered due to the required financial outlay.

Research by \citet{CharltonLaudan2020WebBasedVisualization} shows that a web-based tool specific to MATSim is now possible, given the latest advances in graphic technology embedded in modern web browsers. That tool includes a "front-end" website which runs in the user's browser. Through this site the user organizes model outputs, and defines aggregate visualizations for a few typical use cases. The tool also requires a large, complex "back-end" server component which handles file storage, user authentication, output post-processing, and realtime visualization support. That work is an obvious starting point for this research, but for reasons that will be described below, much of it was ultimately discarded.

@@

Despite two years of development of the above-mentioned client/server solution for other studies, problems in the old system are persistent. The two largest: (1) The massive size of MATSim outputs always preclude loading them into a user's web browser; a single DRT run can be 800Mb or more. This forces post-processing and/or aggregation to happen, which can eliminate many of the benefits of a fully disaggregate analysis. (2) The back-end server design requires extensive effort whenever new visualization capabilities are desired. A strict database design and overwrought user/group management capabilities result in a rock-solid system that is extremely onerous to modify.

It turns out that complex file and user management is unnecessary.\footnote{This invokes the programming mantra "YAGNI" -- You Aren't Gonna Need It: build the simplest possible solution that works, instead of over-engineering things.} From trial and error and iteration emerges a simplified system that eliminates all of the back-end server complexity by eliminating all of the back-end servers:

\begin{itemize}
  \item File storage is delegated to existing servers, instead of creating a special visualization file server. On site, we have an existing "Subversion" server \cite{Collins-SussmanEtc2008SubversionBook} which is adept at large-file storage and version history. All access to the server is via a simple interface that can be easily extended to cloud storage, such as Amazon S3
  \item As the site is intended to be publicly accessible, we discard all user authentication capabilities. If you have a web browser, you can view the results of any runs which are stored \emph{in the public area} of the Subversion server. Naturally some parts of file storage remain off-limits to the greater Internet
  \item Analysts define visualizations by creating small configuration text files in the same folder as model outputs. These files include all of the details of a visualization --- usually just a few lines --- pointing to the input files and configuration settings. This eliminates all the hassle of writing code for user interface input dialog boxes.
  \item Post-processing of outputs must be performed by the analyst as part of their modeling workflow, before uploading to Subversion, since browsers cannot ingest MATSim outputs in their entirety.
\end{itemize}

@@

\begin{itemize}
  \item The overall application framework is based on \textbf{VueJS}, a popular Javascript toolkit for building interactive websites. VueJS is the glue between user interface clicks and the application code that runs underneath.
  \item The agent-based visualizations use custom code ("shaders") written using the \textbf{Deck.gl} visualization library
  \item Older visualizations borrowed from Charlton et al's previous research use the \textbf{Mapbox GL} mapping library
  \item Extensive optimization of memory usage for handling these large datasets with the brilliant \textbf{crossfilter} Javascript library, which manages lightning-fast filtering of millions-of-rows datasets in multiple dimensions
\end{itemize}

The entire codebase is open source and is designed with a "plug-in" architecture so that new visualizations can be developed quickly while relying on common infrastructure for things like file access. To create a new visualization, the developer copies an existing template, modifies the Javascript code to take advantage of the Deck.gl or MapBox libraries as needed, and registers it with the main application. End users can then create visualizations by adding corresponding configuration files to a scenario's file server folder.

@@

The descriptions below fit in the paper's length, but for best understanding of an interactive website, it is better to view the site in situ. The project site remains online and is better demonstrated than described.\footnote{Site is viewable at \href{https://avoev-vsp.github.io}{avoev-vsp.github.io}}

@@

Users can also \emph{compare} the aggregate total mode shift from one scenario to another, in this study comparing DRT scenarios to the base case. These visualizations really highlight the degree to which the new DRT mode is "stealing" trips from existing transit and bicycle modes versus from private vehicles.

@@

Throughout summer 2020, our team iterated on dozens of DRT scenarios: varying pricing structures, presence of drivers vs. fully automated service, service coverage areas, maximum wait time goals, and so on. In August 2020 we began the first public meetings with transit agency staff, with the fully operational website at \href{https://avoev-vsp.github.io}{avoev-vsp.github.io} (available in German only).

@@

Several years of trial and error led us to this specific combination of technologies and techniques. The research team is deeply indebted to the developers of so many components and libraries on which this is based, all of which are freely available online and without which this work would have been entirely impossible. The final product is more than parts merely cobbled together: a usable, useful tool for decision-making now exists, and the DRT-specific extensions serve as a template for further work.

@@

The website remains online at \href{https://avoev-vsp.github.io}{avoev-vsp.github.io}. All code is available at \href{https://github.com/avoev-vsp/avoev-vsp.github.io}{github.com/avoev-vsp} and is licensed under the GNU GPL Version 3 \cite{FSF2007GnuGPL}. The authors welcome feedback and code contributions.
